Bayesian Inference Using Sequential Monte-Carlo Algorithm for Dynamic System Models

### 2020-07-15 MSc-dissertation.tex
New para

\subsection{ABC SMC and model hyperparameters}

@@

@@

\paragraph{Perturbation kernels}

@@

There are several common choice of perturbation kernels. Among them multivariate normal kernel and local M-nearest neighbour model is preferred to be applied on our models.

@@

\chapter{Stuff which no-one will read}

@@

\bibitem{ref:kernel} Filippi, S., Barnes, C. P., Cornebise, J., and Stumpf, M..H. (2013). On optimality of kernels for approximate Bayesian computation using sequential Monte Carlo. Statistical Applications in Genetics and Molecular Biology 12, 1, 87-107

>>>>

\subsection{ABC SMC and model hyperparameters experiments}

@@

[including experiments results and discussions]

@@

\subsubsection{Perturbation kernels}

@@

There are several common choice of perturbation kernels. Among them multivariate normal kernel and local M-nearest neighbour model is preferred to be applied on our models. A covariance matrix $\Sigma_t$ of accepted particles is calculated form previous generation and used in multivariate normal kernel: $K(\theta|\theta_{t-1})\sim\mathcal{N}(\theta_{t-1}, \Sigma_t)$. It is illustrated to be more efficient than uniform kernel and component-wise normal kernel in relecting the true posterior structure \cite{ref:kernel}. It has been proved to perform well in several dynamic system models \cite{ref:abcsysbio, ref:compare, ref:disease} for the parameter estimation and model selection tasks. The $scaling\in(0,1]$ parameter in \verb|pyABC| will be multiplied to the covariance to produce a `narrower distributed' perturbation result.

Local M-NN kernel provided by \verb|pyABC| is also tried to provide a comparison. A local kernel density estimation (KDE) fit is used with M nearest neighbours considered.

The kernel experiments is designed to explore the efficiency of SMC on our dynamical systems. Given the same fixed threshold schedule, kernels that need less total samples and have higher acceptance rate would be our preference. The experiments compared multivariate normal kernel with local MNN kernels with different parameters, using synthetic data to infer back the parameters. The acceptance rate among each time point and total required samples are compared after the experiment to give suggestions on the kernel selection in the real data inference. As the threshold schedule is fixed, the final population should have similar discrepancy to the target data thus here the goodness of fit i.e. recover of target data trends/features and errors of the inferred parameters compared to true values is not discussed.

\subsubsection{Kernel experiment results}

The true value of parameters that are used to generate synthetic data is listed in Appendix B.1. These true values are firstly obtained by fitting a least square fit of model 1 onto the experimental data. The target data [FIGURE HERE] (synthetic data) is then generated using these parameter values via ODE solver in \verb|Python|.

\paragraph{Adaptive functiosn and factors}

\paragraph{Data size, prior range, population size and generations} 1

@@

\chapter{Data and settings}

\section{Infer-back experiments}

@@

\bibitem{ref:kernel} Filippi, S., Barnes, C. P., Cornebise, J., and Stumpf, M..H. (2013). On optimality of kernels for approximate Bayesian computation using sequential Monte Carlo. Statistical Applications in Genetics and Molecular Biology 12, 1, 87-107.

\bibitem{ref:abcsysbio} Liepe, J., Kirk, P., Filippi, S., Toni, T., Barnes, C.P., and Stumpf, M.P.H., 2014. A framework for parameter estimation and model selection from experimental data in systems biology using approximate Bayesian computation. Nature Protocols. 9 (2) pp. 439–456.

\bibitem{ref:compare} Daly AC, Cooper J, GavaghanDJ, Holmes C. 2017. Comparing two sequentialMonte Carlo samplers for exact andapproximate Bayesian inference on biologicalmodels.J. R. Soc. Interface14: 20170340.

\bibitem{ref:disease} Minter, A. and Retkute, R., 2019. Approximate Bayesian Computation for infectious disease modelling. Epidemics, 29, p.100368.
